On semidefinite-representable sets over valued fields

### 2025-12-23 issac_biblatex_tentative/paper_CNV.tex
commit by Simone

\begin{example}
\label{fermat_quartic}
The basic closed semialgebraic set $\mathcal{R} = \{(x_1,x_2) \in \R^2 : 1-x_1^4-x_2^4 \geq 0\}$ is not a spectrahedron but it is semidefinite representable with representation:
$$
%%\mathcal{R} =
\left\{\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} \in \R^2 :
\exists\,
\begin{bmatrix} y_1 \\ y_2 \end{bmatrix} \in \R^2, \,
%\left[
%  \begin{smallmatrix}
%    1+y_1 & y_2   &   0 &   0 & 0 & 0 \\
%      y_2 & 1-y_1 &   0 &   0 & 0 & 0 \\
%        0 &     0 &   1 & x_1 & 0 & 0 \\
%        0 &     0 & x_1 & y_1 & 0 & 0 \\
%        0 &     0 &   0 &   0 & 1 & x_2 \\
%        0 &     0 &   0 &   0 & x_2 & y_2
%  \end{smallmatrix}
%  \right]
\diag\left(
\left[
\begin{smallmatrix}
  1+y_1 & y_2 \\
  y_2 & 1-y_1
\end{smallmatrix}
\right],
\left[
\begin{smallmatrix}
  1 & x_1 \\
  x_1 & y_1
\end{smallmatrix}
\right],
\left[
\begin{smallmatrix}
  1 & x_2 \\
  x_2 & y_2
\end{smallmatrix}
\right]
\right)
\succeq 0
\right\}
$$
\end{example}

@@

In this section, we prove that the class of polyhedra is stable
by applying affine maps:

\begin{theorem}[DI] 
Let $f : K^n \to K^m$ be an affine map and $\PP$ a polyhedron in $K^n.$
Then $f(\PP)$ is a polyhedron. \label{theo:DI}
\end{theorem}

@@

\begin{figure}[!ht] 
\centering
\scalebox{0.5}{
\begin{tikzpicture}[
  node distance=1cm and 1cm,
  box/.style={draw, rounded corners, minimum width=2.2cm, minimum height=1cm, align=center},
  thickarrow/.style={->, >=Stealth, line width=1.2pt}
]
% Nodes
\node[box] (DIdiag) {DI${}_{diag}$};
\node[box, right=of DIdiag] (DIproj) {DI${}_{proj}$};
\node[box, right=of DIproj] (DIl) {DI${}_{l}$};
\node[box, right=of DIl] (DIi) {DI${}_{i}$};
\node[box, right=of DIi] (DI) {DI};
\node[box, above=of DIproj] (DIimm) {DI${}_{imm}$};
\node[box, below=of DIproj] (DIaut) {DI${}_{aut}$};
% Arrows
\draw[thickarrow] (DIdiag) -- (DIproj);
\draw[thickarrow] (DIproj) -- (DIl);
\draw[thickarrow] (DIl) -- (DIi);
\draw[thickarrow] (DIi) -- (DI);
\draw[thickarrow] (DIimm) -- (DIl);
\draw[thickarrow] (DIaut) -- (DIl);
\end{tikzpicture}
}
\caption{Strategy of proof of Theorem \ref{theo:DI}.}
\label{fig:di-properties}
\end{figure}

We begin with the case of linear maps in $\GL_n(O_K)$.

\begin{lemma}[DI${}_{aut}$]
Let $f \in GL_n(O_K)$ and $\PP$ a polyhedron in $K^n$.
Then $f(\PP)$ is a polyhedron. \label{lem:DIaut}
If in addition, $\PP$ is defined only by inequalities,
then so is $f(\PP)$.
\end{lemma}
\begin{proof}
If $\PP$ is defined (in matrix form) by \[\left\lbrace X \in K^n,  AX+B \geq 0, CX+D=0 \right\rbrace\] and $U$ is the matrix of $f$ in the canonical basis of $K^n$, 
let $\Qc$ be defined by \[\left\lbrace Y \in K^n,  AU^{-1} Y+B \geq 0, CU^{-1} Y+D=0 \right\rbrace.\]
Then it is clear that $\Qc$ is a polyhedron and $f(\PP)=\Qc.$
It is also clear that if $\PP$ is defined only by inequalities,
then so is $f(\PP)$.
\end{proof}

@@

\begin{lemma}[DI${}_{diag}$]
Let $f : K^n \to K^n$ be a linear map and $\PP$ a polyhedron in $K^n$ defined only by inequalities. We assume that the matrix of $f$ in the canonical basis of $K^n$ is diagonal.
Then $f(\PP)$ is a polyhedron. \label{lem:DIdiag}
\end{lemma}
\begin{proof}
As in the proof of Lemma \ref{lem:DIaut} DI${}_{aut}$, a trivial change
of variables is enough to conclude.
\end{proof}

We now treat the case where $f$ is a projection erasing the last coordinate.

\begin{lemma}[DI${}_{proj}$]
Let $f : K^n \to K^{n-1}, \: (x_1,\dots,x_{n-1},x_n) \mapsto (x_1,\dots,x_{n-1})$ and $\PP$ a polyhedron in $K^n$ defined only by inequalities. 
Then $f(\PP)$ is a polyhedron defined only by inequalities. \label{lem:DIproj}
\end{lemma}
\begin{proof}
We assume that $\PP \neq \emptyset$.
Let us write the conditions defining $\PP$
in matrix form:
$\val(A X+B) \geq 0$ 
with $A \in M_{q,n}(K)$, $B \in K^q$
and $X=(x_1,\dots,x_n)^\intercal$.

We can compute a Smith Normal Form
of the $q \times (n-1)$ matrix defined by the 
$n-1$ first columns of $A$
to obtain a decomposition of the form:
\[ Q_A A P_A^{-1} = \begin{bmatrix}
\delta_1	& 		& 			&   &		 &  &a_1	\\
			& \ddots& 			& 0	&		 &	&		\\
			&		& \delta_l  &   & 		 &	&		\\
			&		&			&0  & 		 &	& 		\\
			&		&			&   & \ddots &	&		\\
			&		&			&	&		 & 0&a_{n-1} \\
			&		&			&	&		 &	&a_n 	\\
			&		&			&	&		 &	&    	\\
			&		&			&	&		 &	&a_q 	\\			
\end{bmatrix},\]
with $Q_A \in \GL_q(\OK)$ and $P_A \in \GL_n(\OK)$ such that $P_A(e_n)=e_n.$

>>>>

\begin{example}
\label{fermat_quartic}
The basic closed semialgebraic set $\mathcal{R} = \{(x_1,x_2) \in \R^2 : 1-x_1^4-x_2^4 \geq 0\}$ is not a spectrahedron but it is a spectrahedral shadow with semidefinite representation
$$
%%\mathcal{R} =
\left\{x \in \R^2 :
%%\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} \in \R^2 :
\exists\, y \in \R^2,
%%\begin{bmatrix} y_1 \\ y_2 \end{bmatrix} \in \R^2, \,
%\left[
%  \begin{smallmatrix}
%    1+y_1 & y_2   &   0 &   0 & 0 & 0 \\
%      y_2 & 1-y_1 &   0 &   0 & 0 & 0 \\
%        0 &     0 &   1 & x_1 & 0 & 0 \\
%        0 &     0 & x_1 & y_1 & 0 & 0 \\
%        0 &     0 &   0 &   0 & 1 & x_2 \\
%        0 &     0 &   0 &   0 & x_2 & y_2
%  \end{smallmatrix}
%  \right]
\diag\left(
\left[
\begin{smallmatrix}
  1+y_1 & y_2 \\
  y_2 & 1-y_1
\end{smallmatrix}
\right],
\left[
\begin{smallmatrix}
  1 & x_1 \\
  x_1 & y_1
\end{smallmatrix}
\right],
\left[
\begin{smallmatrix}
  1 & x_2 \\
  x_2 & y_2
\end{smallmatrix}
\right]
\right)
\succeq 0
\right\}
$$
\end{example}

@@

In this section, we prove that the class of polyhedra is stable
under application of affine maps:

\begin{theorem}[direct image (DI)] 
Let $f : K^n \to K^m$ be an affine map and $\PP$ a polyhedron in $K^n.$
Then $f(\PP)$ is a polyhedron. \label{theo:DI}
\end{theorem}

@@

\begin{figure}[!ht] 
\centering
\scalebox{0.55}{
\begin{tikzpicture}[
  node distance=1cm and 1cm,
  box/.style={draw, rounded corners, minimum width=2.2cm, minimum height=1cm, align=center},
  thickarrow/.style={->, >=Stealth, line width=1.2pt}
]
% Nodes
\node[box] (DIdiag) {DI${}_{diag}$};
\node[box, right=of DIdiag] (DIproj) {DI${}_{proj}$};
\node[box, right=of DIproj] (DIl) {DI${}_{l}$};
\node[box, right=of DIl] (DIi) {DI${}_{i}$};
\node[box, right=of DIi] (DI) {DI};
\node[box, above=of DIproj] (DIimm) {DI${}_{imm}$};
\node[box, below=of DIproj] (DIaut) {DI${}_{aut}$};
% Arrows
\draw[thickarrow] (DIdiag) -- (DIproj);
\draw[thickarrow] (DIproj) -- (DIl);
\draw[thickarrow] (DIl) -- (DIi);
\draw[thickarrow] (DIi) -- (DI);
\draw[thickarrow] (DIimm) -- (DIl);
\draw[thickarrow] (DIaut) -- (DIl);
\end{tikzpicture}
}
\caption{Strategy of proof of Theorem \ref{theo:DI}.}
\label{fig:di-properties}
\end{figure}

We begin with the case of linear maps in $\GL_n(O_K)$.
We often fix a linear basis $e_1,\ldots,e_n$ of $K^n$ which we call canonical\footnote{j'ai rajouté ça}.

\begin{lemma}[DI${}_{aut}$]
Let $f \in GL_n(O_K)$ and $\PP$ a polyhedron in $K^n$.
Then $f(\PP)$ is a polyhedron. \label{lem:DIaut}
If in addition, $\PP$ is defined only by inequalities,
then so is $f(\PP)$.
\end{lemma}
\begin{proof}
  If $\PP$ is defined (in matrix form) by
  \[
  \PP=\left\lbrace x \in K^n,  Ax+B \geq 0, Cx+D=0 \right\rbrace
  \] and $U$ is the matrix of
  $f$ in the canonical basis of $K^n$, 
  let
  \[\Qc=\left\lbrace y \in K^n,  AU^{-1} y+B \geq 0, CU^{-1} y+D=0 \right\rbrace.\]
  Then it follow that $\Qc$ is a polyhedron and $f(\PP)=\Qc.$
  It is also clear that if $\PP$ is defined only by inequalities,
  then so is $f(\PP)$.
\end{proof}

@@

\begin{lemma}[DI${}_{diag}$]
  Let $f : K^n \to K^n$ be a linear map and $\PP$ a polyhedron in $K^n$ defined only by inequalities.
  If the matrix of $f$ in the canonical basis of $K^n$ is diagonal,
  then $f(\PP)$ is a polyhedron. \label{lem:DIdiag}
\end{lemma}
\begin{proof}
As in Lemma \ref{lem:DIaut}, a change
of variables is enough to conclude.
\end{proof}

We assume now $f$ is a projection erasing the last coordinate.

\begin{lemma}[DI${}_{proj}$]
  Let $f : K^n \to K^{n-1}, \: (x_1,\dots,x_{n-1},x_n) \mapsto (x_1,\dots,x_{n-1})$, and let
  $\PP$ be a polyhedron in $K^n$ defined only by inequalities. 
  Then $f(\PP)$ is a polyhedron defined only by inequalities. \label{lem:DIproj}
\end{lemma}
\begin{proof}
We assume that $\PP \neq \emptyset$.
Let us write the conditions defining $\PP$
in matrix form:
$\val(A x+B) \geq 0$ 
with $A \in M_{q,n}(K)$, $B \in K^q$,
and $x=(x_1,\dots,x_n)$ a column vector.

We can compute a Smith Normal Form SNF($A'$)
of the $q \times (n-1)$ matrix $A'$ defined by the 
$n-1$ first columns of $A$
\[
\mathrm{SNF}(A') = \diag(\delta_1,\ldots,\delta_r,{\bf 0}_{q-r,n-1-r})
\]

to obtain a decomposition of the form:
\[
Q_A A P_A^{-1} =
\begin{bmatrix}
  &                  & & a_1 \\
  & \mathrm{SNF}(A') & & \vdots \\
  &                  & & a_{n-1} \\
  &                  & &
\end{bmatrix}
%\begin{bmatrix}
%  \delta_1	& 		& 			&   &		 &  &a_1	\\
%  & \ddots& 			& 0	&		 &	&		\\
%  &		& \delta_l  &   & 		 &	&		\\
%  &		&			&0  & 		 &	& 		\\
%  &		&			&   & \ddots &	&		\\
%  &		&			&	&		 & 0&a_{n-1} \\
%  &		&			&	&		 &	&a_n 	\\
%  &		&			&	&		 &	&    	\\
%  &		&			&	&		 &	&a_q 	\\			
%\end{bmatrix},
\]
with $Q_A \in \GL_q(\OK)$ and $P_A \in \GL_n(\OK)$ such that $P_A(e_n)=e_n.$
